Multiple Output Regression

### 2014-02-13 paper/paper.tex
discussion on learning of inducing inputs

>>>>

\subsection{Analysis on Learning of the Inducing Inputs}
In this section we present some analysis on learning of the inducing inputs under the stochastic variational inference procedure on some toy problems.
We use three real-valued truth functions of scalar inputs: the first one was used in Snelson et al, the second one is a function $f(x) = \sin(x) + \cos(x)$, and the third one is a sample function generated from a GP with squared exponential with ARD covariance function with a lengthscale of 2 and signal variance of 1.

\noindent \paragraph{Experimental Settings} We use gpsvi to optimize for \textit{all} of the parameters in the model, i.e. the variational parameters of the posterior, the covariance hyperparameters, the noise hyperparameter, and the inducing inputs.
For the hyperparameters, we use a momentum of $0.9$ and a fixed learn rate of $1e-5$. 
For the variational parameters, we use a learn rate of $0.01$ (no momentum was used).
We vary the learn rate of the inducing inputs from $1e{-2}$ to $1e{-6}$.
Using momentum for the inducing inputs does not seem to have much effects.
The maximum number of iterations is 500, the batch size 5, and the number of inducing values 10.

% effect of learn rate on inducing inputs
\noindent \paragraph{Effects of the Learn Rates }
In Figure \ref{fig1}, \ref{fig2}, and \ref{fig3}, we show the lower bounds, predictive distributions, and the learned inducing inputs using stochastic variational inference with varying learn rates for the inducing inputs.
We also show the results with FITC for comparison.
It can be seen from the figures that, as typical in stochastic optimization, the behavior of GPSVI is very sensitive to the learn rate.
In particular, when the rate is slow e.g. $1e{-5}$, almost no progress was made in learning the inducing inputs.
However, when the rate increases to e.g. $1e{-3}$, the inducing inputs are spread out more evenly in the input space compared to the initial locations.
Although not shown the in figures, when the learn rate is too large, e.g. $1e{-02}$, the inducing inputs may spread to regions far outside of the training intervals.
An example of this phenomenon can be observed in the top left figure in Figure \ref{fig1}.
The learn rates also affect the evidence lower bound which seems to wiggle more compared to not learning the inducing inputs.
The predictive distributions are sensible particularly because the toy functions exhibit sparsity and are easy to learn.  \\

\noindent The results of this qualitative analysis do not seem to suggest that learning of the inducing inputs does not work in GPSVI.
Its effectiveness perhaps rests upon empirical performance on real datasets.

\noindent \paragraph{The Noise Parameter} It seems harder to set the initial value and the learning rate of the noise parameter compared to standard GP. 
While a large range of values of noise can be used in standard gp or fitc (e.g. 0.5), a very small value of noise $1e{-02}$, which is the true noise, is required for gpsvi to work well.
It appears that an adaptive learning rate may be required for the noise parameter.

\begin{figure*}
\centering
\begin{tabular}{ccc}
\includegraphics[scale=0.3]{figures/func1-svi-lrate1e-03.eps} &
\includegraphics[scale=0.3]{figures/func1-svi-lrate1e-04.eps} &
\includegraphics[scale=0.3]{figures/func1-svi-lrate1e-05.eps} \\
\includegraphics[scale=0.3]{figures/func1-svi-lrate1e-03-bound.eps} &
\includegraphics[scale=0.3]{figures/func1-svi-lrate1e-04-bound.eps} &
\includegraphics[scale=0.3]{figures/func1-svi-lrate1e-05-bound.eps} \\ 
(a) learn rate = $1e{-3}$ & (b) learn rate = $1e{-4}$ & (c) learn rate = $1e{-5}$ \\
\multicolumn{3}{c}{\includegraphics[scale=0.4]{figures/func1-fitc.eps}}
\end{tabular}
\caption{Predictive distributions and learned inducing inputs by GPSVI and FITC for the first function. Top row: GPSVI with different learning rates (the rates decreases from left to right). Bottom row: FITC. Magenta dots : training points. Solid blue line: predictive mean. Grey-shaded area and solid black lines: two standard deviations of the predictive distributions. Black (+) crosses: initial locations of inducing inputs. Magenta (+) crosses: learned locations of inducing inputs.  Middle row: the evidence lower bound of gpsvi vs. iteration.}
\label{fig1}
\end{figure*}

\begin{figure*}
\centering
\begin{tabular}{ccc}
\includegraphics[scale=0.3]{figures/func2-svi-lrate1e-03.eps} &
\includegraphics[scale=0.3]{figures/func2-svi-lrate1e-04.eps} &
\includegraphics[scale=0.3]{figures/func2-svi-lrate1e-05.eps} \\
\includegraphics[scale=0.3]{figures/func2-svi-lrate1e-03-bound.eps} &
\includegraphics[scale=0.3]{figures/func2-svi-lrate1e-04-bound.eps} &
\includegraphics[scale=0.3]{figures/func2-svi-lrate1e-05-bound.eps} \\ 
(a) learn rate = $1e{-3}$ & (b) learn rate = $1e{-4}$ & (c) learn rate = $1e{-5}$ \\
\multicolumn{3}{c}{\includegraphics[scale=0.4]{figures/func2-fitc.eps}}
\end{tabular}
\caption{Predictive distributions and learned inducing inputs by GPSVI and FITC for the second function. The legends are same as in Figure 1.}
\label{fig2}
\end{figure*}

\begin{figure*}
\centering
\begin{tabular}{ccc}
\includegraphics[scale=0.3]{figures/func3-svi-lrate1e-03.eps} &
\includegraphics[scale=0.3]{figures/func3-svi-lrate1e-04.eps} &
\includegraphics[scale=0.3]{figures/func3-svi-lrate1e-05.eps} \\
\includegraphics[scale=0.3]{figures/func3-svi-lrate1e-03-bound.eps} &
\includegraphics[scale=0.3]{figures/func3-svi-lrate1e-04-bound.eps} &
\includegraphics[scale=0.3]{figures/func3-svi-lrate1e-05-bound.eps} \\ 
(a) learn rate = $1e{-3}$ & (b) learn rate = $1e{-4}$ & (c) learn rate = $1e{-5}$ \\
\multicolumn{3}{c}{\includegraphics[scale=0.4]{figures/func3-fitc.eps}}
\end{tabular}
\caption{Predictive distributions and learned inducing inputs by GPSVI and FITC for the third function. The legends are same as in Figure 1.}
\label{fig3}
\end{figure*}
